---
title: "p8105_hw3_dk2759"
author: "Darwin Keung"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup}

library(tidyverse)
library(ggridges)
library(dplyr)

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = 1.0,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```


Homework 3
Context
This assignment reinforces ideas in Visualization and EDA.

Due date
Due: October 15 at 4:00pm.

Points
Problem 0: 10 points
Problem 1: 30 points
Problem 2: 30 points
Problem 3: 30 points
Problem 0
This “problem” focuses on structure of your submission, especially the use git and GitHub for reproducibility, R Projects to organize your work, R Markdown to write reproducible reports, relative paths to load data from local files, and reasonable naming structures for your files.

To that end:

create a public GitHub repo + local R Project; we suggest naming this repo / directory p8105_hw3_YOURUNI (e.g.  p8105_hw3_ajg2202 for Jeff), but that’s not required
create a single .Rmd file named p8105_hw3_YOURUNI.Rmd that renders to github_document
Your solutions to Problems 1, 2, and 3 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems.

For this Problem, we will assess adherence to the instructions above regarding repo structure, git commit history, and whether we are able to knit your .Rmd to ensure that your work is reproducible. Adherence to appropriate styling and clarity of code will be assessed in Problems 1+ using the homework style rubric.

This homework includes figures; the readability of your embedded plots (e.g. font sizes, axis labels, titles) will be assessed in Problems 1+.

## Problem 1
This problem uses the BRFSS data. DO NOT include this dataset in your local data directory; instead, load the data from the  p8105.datasets package.

```{r}
library(p8105.datasets)
```

#### Read and clean BRFSS data

Load the BRFSS data from the `p8105.datasets` package.

```{r read_brfss}
data(brfss_smart2010)
```

Cleaning
* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Excellent” to “Poor”

```{r clean_brfss}
brfss_smart2010_factor =
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%
  janitor::clean_names()  %>%
  arrange(response)
``` 

#### Questions

In 2002, which states were observed at 7 locations?
ANS: CT, FL, NC have 7 distinct observation locations.

```{r 7_locations}
brfss_smart2010_factor %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  distinct(locationdesc) %>% 
  count() %>% 
  filter(n == 7)
```

Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r spaghetti_plot}
brfss_smart2010_plot = brfss_smart2010_factor %>% 
  group_by(locationabbr, year) %>% 
  distinct(locationdesc) %>%
  summarize(n_obs = n())

ggplot(brfss_smart2010_plot, aes(x = year, y = n_obs)) + 
  geom_line(aes(color = locationabbr), alpha = .5) +
  labs(
    title = "BRFSS Observation Locations by State from 2002 to 2010",
    x = "Year",
    y = "Number of Locations",
    caption = "Data from the BRFSS package"
  ) + 
  viridis::scale_color_viridis(
    name = "State", 
    discrete = TRUE
  )
```

Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

This table below shows the proportion of "Excellent" responses across all of NY State. The average was .240 in 2002 the decreased to .225 in 2006, and rose slightly to .227 in 2010.

```{r table_of_excellence}
brfss_smart2010_factor %>% 
  filter(locationabbr == "NY", year %in% c(2002, 2006, 2010),response == "Excellent") %>% 
  group_by(year) %>%
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>%
  summarize(mean_excellent = mean(excellent), 
            sd_excellent = sd(excellent, na.rm = TRUE)) %>% 
  mutate(mean_excellent = 
           mean_excellent / 100, sd_excellent = sd_excellent / 100) %>% 
  knitr::kable(digits = 3)
```

For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

This five-panel plot shows that the highest average response rates were very good. People tend to be kinder as the top three responses are excellent, very good, good. Whereas fair and poor scores were rarer.

```{r average_response_category}
brfss_smart2010_factor %>% 
  group_by(year, locationabbr, response) %>%
  summarize(mean_response = mean(data_value/100)) %>% 
  ggplot(aes(x = year, y = mean_response, color = locationabbr)) + 
  geom_line(alpha = 0.4) + 
  facet_grid(~response) +
  labs(
    title = "BRFSS Average Response by State from 2002 to 2010",
    x = "Year",
    y = "Average Response",
    caption = "Data from the BRFSS package"
  ) + 
  viridis::scale_color_viridis(
    name = "State", 
    discrete = TRUE
  )
```

## Problem 2

#### Read and clean Instacart data

```{r}
data(instacart)
```

This problem uses the Instacart data. DO NOT include this dataset in your local data directory; instead, load the data from the  p8105.datasets package (it’s called instacart).

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations. Then, do or answer the following (commenting on the results of each):

#### Summarize

This dataset has `r nrow(instacart)` observations and `r ncol(instacart)` variables. Other than `eval_set`, `product_name`, `aisle`, and `department`, all variables are class `integer`. There are **`r instacart %>% distinct(aisle_id) %>% count()` aisles** and **`r instacart %>% distinct(department_id) %>% count()` departments** selling a total of **`r instacart %>% distinct(product_name) %>%  count()` unique products**

#### Questions

How many aisles are there, and which aisles are the most items ordered from?
There are **`r instacart %>% distinct(aisle_id) %>% count()` aisles**. `Fresh vegetables` and `fresh fruits` have the most ordered items.

```{r aisle_order_counts}
instacart %>%
  group_by(aisle) %>% 
  count(aisle) %>%
  rename(count_aisle = n) %>% 
  arrange(desc(count_aisle)) %>%
  head() %>% 
  knitr::kable()
```

Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

The code below plots number of items order in each aisle ordered by `aisle_id`. People really really like to order fruits and veggies over other items by far.

```{r instacart_orders_plot}
instacart_plot = instacart %>%
  group_by(aisle, aisle_id) %>% 
  count(aisle) %>%
  rename(count_aisle = n) %>%
  arrange(desc(aisle_id))

ggplot(instacart_plot, aes(x = aisle, y = count_aisle)) + 
  geom_point(aes()) +
  labs(
    title = "Instacart Items Ordered per Aisle",
    x = "Aisle",
    y = "Number of Orders",
    caption = "Data from the Instacart package"
  ) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

The code below produces this table. From baking ingredients: zero calorie sweetener; dog food care: zero grain turkey dog food; packaged vegetables fruits: Zucchini Squash, Baby Courgette.
```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  summarize(popular_item = max(product_name)) %>% 
  head() %>% 
  knitr::kable()
```
  
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

The latest average order time for both Coffe Ice Cream and Pink Lady Apples occurs on Tuesday afternoon.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream"))  %>%
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>% 
  rename("Product Name" = "product_name", "Saturday" = "0", 
         "Sunday" = "1", "Monday" = "2", "Tuesday" = "3", 
         "Wednesday" = "4", "Thursday" = "5", "Friday" = "6") %>%
  knitr::kable(digits = 3)
```

## Problem 3
This problem uses the NY NOAA data.

#### Summary

```{r ny_noaa}
data(ny_noaa)
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

```{r ny_noaa_summary}
skimr::skim(ny_noaa)
n_missing_tmin = ny_noaa %>%
  mutate(tmin_na = is.na(tmin)) %>%
  count(tmin_na) %>% filter(tmin_na == TRUE)

n_missing_prcp = ny_noaa %>%
  mutate(prcp_na = is.na(prcp)) %>%
  count(prcp_na) %>% filter(prcp_na == TRUE)
```


This dataset has `r nrow(ny_noaa)` observations and `r ncol(ny_noaa)` variables. This dataset contains readings from `r ny_noaa %>% summarize(n_distinct(id))` weather stations. There are the following  variables. 

* Character: `id`, `tmax`, `tmin`. 
* Date: `date`
* Integer: `prcp`, `snow`, `snwd`

id: Weather station ID
date: Date of observation
prcp: Precipitation (tenths of mm)
snow: Snowfall (mm)
snwd: Snow depth (mm)
tmax: Maximum temperature (tenths of degrees C)
tmin: Minimum temperature (tenths of degrees C)

There is a lot of missing data for `tmax` and `tmin`, as well `prcp`, `snow`, `snwd`. Only  `r ny_noaa %>% na.omit() %>% nrow()` observations have no missing values. For the `tmin` variable there are `r n_miss_tmin$n` missing values and for the `pcrp` variable there are `r n_missing_prcp$n` missing values.

#### Cleaning Data

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

```{r}
ny_noaa =  separate(date, into = c("year", "month", "day"))
```


Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.